---
title: "Calculating Shannon Diversity"
format: html
editor: visual
---

## Set up

```{r}
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("vegan")
#install.packages("brms")
#install.packages("tidyverse")
#install.packages("readxl")
#install.packages("usethis")
#install.packages("rmarkdown")
#install.packages("glmmTMB")
#install.packages("Rtools")
library(dplyr)
library(ggplot2)
library(vegan)
library(brms)
library(tidyverse)
library(readxl)
library(usethis)
library(rmarkdown)
library(glmmTMB)
library(lme4)


df <- read.csv("DATA collection_D_Gutierrez - Day1.csv")
head(df)
```

## Connecting github

```{r}
#setting to your user
#use_git_config(user.name = "tintin969")
#used to check what is connected
#git_sitrep()
#creates a new token to connect to github
#usethis::create_github_token()
#setting which is connected with token code
#gitcreds::gitcreds_set()
#checking that it worked
#git_sitrep()
```

# numeric

1.  Make all numeric columns numeric
2.  make all NA's into 0

```{r}
df4<-df |> mutate(across(c("Seeds.amt.eaten..g.","Neotoma", "Peromyscus", "Dipodomys", "Chaetodipus", "Otospermophilus", "Reithrodonomys", "Sylvilagus", "Bird", "Microtus", "other", "Raccoon"), as.numeric))

is.numeric(df4$Seeds.amt.eaten..g.)
df4 <- df4 %>% mutate(across(where(is.numeric), ~ifelse(is.na(.x), 0, .x)))


```

# Fixing condition columns

1.  duplicating the column

    ```{r}
    df4$Condition_Copy <- df4$Condition..L.D.
    ```

2.  replace numbers with nothing

    ```{r}
    df4$Condition_Copy <- gsub("[0-9]", "", df4$Condition..L.D.)

    #head(df$con)

    ```

# Shannon diversity

```{r}

#?diversity

#calculate shannon diversity for each row



#?mutate
colnames(df4)

#Select for species to calc
df5 <- df4 |>  select(c("Neotoma", "Dipodomys", "Reithrodonomys","Microtus","Chaetodipus","Sylvilagus","Otospermophilus","Bird","other", "Raccoon","Peromyscus"))

#try to calculate for diversity
df4$Shannon.Diversity <- df5 |> diversity(index = "shannon")

#dpmt work :(
#df_shdv <- df4 |> mutate(Shannon.Diversity == diversity(df5, index = "shannon"))

head(df4)
df4 <- df4 %>% filter(!row_number() %in% c(17, 34, 51, 68, 85, 102, 119, 136, 153, 170, 187, 204)
```

# Graph

```{r}
#generate graph
df4 |> ggplot(mapping = aes(x = Seed.amt.after..g., y = Shannon.Diversity, color = Condition_Copy)) + geom_point()+geom_jitter()
#the animals tend to eat all the seed therefore it is a 0 inflat
```

```{r}
## observations vs light (peromysus)
df4 |> ggplot(mapping = aes(x = Condition_Copy, y = Peromyscus, color = Condition_Copy)) + geom_boxplot() + scale_y_log10()
```

```{r}
# observations vs light dipodymys
df4 |> ggplot(mapping = aes(x = Condition_Copy, y = Dipodomys, color = Condition_Copy)) + geom_boxplot() + scale_y_log10()
```

```{r}
#observaitons vs ligth chaetodipus
df4 |> ggplot(mapping = aes(x = Condition_Copy, y = Chaetodipus, color = Condition_Copy)) + geom_boxplot() + scale_y_log10()
```

```{r}
#seed amount eaten vs condition
df4 |> ggplot(mapping = aes(x = Condition_Copy, y = Seeds.amt.eaten..g., color = Condition_Copy)) + geom_boxplot() + scale_y_log10()
```

```{r}
# visits vs condition
df4 |> ggplot(mapping = aes(x = Condition_Copy, y =  cbind(Peromyscus,Dipodomys,Chaetodipus) , color = Condition_Copy)) + geom_boxplot() + scale_y_log10()
```

# Linear Mixed-effects model

linear mixed effects model:

generalized linear mixed model gamma distribution with log link for positive continuous GUD

beta distrubution if modeled as proportion of seed remaining

0 inflated glmms use glmmTMB

# GUD

Giving up density (GUD) = Seed.amt.after..g.

light = Condition_Copy

Microhabitat = ? not relevant

```{r}
R
library(lme4)
library(lmerTest) # For p-values via Satterthwaite approximation
library(DHARMa)   # For residual diagnostics

# 1. Fit the Linear Mixed-Effects Model
# Note: Log-transform GUD if residuals exhibit right-skew or non-constant variance
gud_model <- lmer(
  log(GUD + 1) ~ Light * Microhabitat + Temp_C + (1 | Station) + (1 | Night),
  data = foraging_data
)

# 2. Model Summary and Type III ANOVA
summary(gud_model)
anova(gud_model, type = "3")

# 3. Residual Diagnostics
residuals_sim <- simulateResiduals(fittedModel = gud_model)
plot(residuals_sim)


```

```{r}
m.test <- 
  brm(data = df4, # Give the model the pie_crab data
      # Choose a gaussian (normal) distribution
      family = gaussian,
      # Specify the model here. 
      Seed.amt.after..g. ~ Condition_Copy + 1,
      # Here's where you specify parameters for executing the Markov chains
      # We're using similar to the defaults, except we set cores to 4 so the analysis runs faster than the default of 1
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      # Setting the "seed" determines which random numbers will get sampled.
      # In this case, it makes the randomness of the Markov chain runs reproducible 
      # (so that both of us get the exact same results when running the model)
      seed = 4,
      # Save the fitted model object as output - helpful for reloading in the output later
      file = "output/m.test2")
```
